First steps towards a formalization of Forcing

### 2018-06-24 LSFA/intro.tex
cambios último momento

It can be shown elementally that if $M$ and $N$ are transitive,
$\phi^N$ holds if and only if $\phi^M$ holds,  for $x,y,z\in N$. We
say then that $\phi$ is \emph{absolute between $N$ and $M$.}
%% MIGUEL: "truth of axioms in $M$" significa que $M$ satisface los
%% axiomas?
The concepts of relativization and absoluteness are central to the
task of transferring truth of axioms in $M$ to $L^M$, and constitute
the hardest part of Paulson's development.

@@

%% MIGUEL: redacción alternativa del párrafo siguiente.
Every  $y \in M[G]$ is obtained from an element $\dot y$ of $M$, thus
elements of $M$ are construed as \emph{names} or codes for elements of
$M[G]$.
The decoding is given by the function
$\val$ which takes the generic filter $G$ as a parameter. To
prove that $M$ is contained in $M[G]$ it suffices to give a name for
each element of $M$; we define the function $\chk$ which assigns
a name for each $x\in M$. Showing that $\chk(x)\in M$
when $x\in M$ involves some technical issues that will
be addressed in a further work. We explain names, $\val$, and
$\chk$ in Section~\ref{sec:names}.

>>>>

It can be shown elementarily that if $M$ and $N$ are transitive,
$\phi^N$ holds if and only if $\phi^M$ holds,  for $x,y,z\in N$. We
say then that $\phi$ is \emph{absolute between $N$ and $M$.}
%% MIGUEL: "truth of axioms in $M$" significa que $M$ satisface los
%% axiomas?
The concepts of relativization and absoluteness are central to the
task of transferring truth of axioms in $M$ to $L^M$, and constitute
the hardest part of Paulson's development.

@@

%% MIGUEL: redacción alternativa del párrafo siguiente.
Every  $y \in M[G]$ is obtained from an element $\dot y$ of $M$, thus
elements of $M$ are construed as \emph{names} or codes for elements of
$M[G]$.
The decoding is given by the function
$\val$, which takes the generic filter $G$ as a parameter. To
prove that $M$ is contained in $M[G]$ it suffices to give a name for
each element of $M$; we define the function $\chk$ which assigns
a name for each $x\in M$. Showing that $\chk(x)\in M$
when $x\in M$ involves some technical issues that will
be addressed in a further work. We explain names, $\val$, and
$\chk$ in Section~\ref{sec:names}.

### 2018-06-24 LSFA/names.tex
cambios último momento

On the one hand, the only requirement on the set  \isatt{G} is that it 
contains the top \isa{one} of the poset \isa{P}. On the other hand, it
is necessary that all the relevant names are indeed in $M$. It
requires a serious development  to fulfill this assumption. One of the
hardest parts of Paulson's formalization of constructibility involves
showing that models are closed under recursive construction. We will
eventually formalize the that if $M\models\ZFC$ and the arguments of
\isa{wfrec} are in $M$, then its value also is. This will require
to adapt to models $M$ several locales defined in \cite{MR2507047}
that were intended to be used for the class of constructible sets.

>>>>

On the one hand, the only requirement on the set  \isatt{G} is that it 
contains the top \isa{one} of the poset \isa{P}. On the other hand, it
is necessary that all the relevant names are indeed in $M$. It
requires a serious development  to fulfill this assumption. One of the
hardest parts of Paulson's formalization of constructibility involves
showing that models are closed under recursive construction. We will
eventually formalize the that if $M\models\ZFC$ and the arguments of
\isa{wfrec} are in $M$, then its value also is. This will require
to adapt to models $M$ several locales defined in \cite{paulson_2003}
that were intended to be used for the class of constructible sets.
